---
title: "Bets Optimization"
author: "Michael Cao"
date: "2024-03-31"
output: html_document
---

```{r setup, include=FALSE, warning=F}
knitr::opts_chunk$set(echo = TRUE)
library(lpSolve)
library(plotly)
library(gtools)
library(data.table)
default <- options()
```

```{r, message=F}
download.file("https://www.football-data.co.uk/mmz4281/2324/E0.csv", 
              "Betting Odds/E0 (2023-24).csv") # manual-refresh
bets <- fread('Betting Odds/E0 (2023-24).csv')
write.csv(bets, "Betting Odds/E0 (2023-24).csv")
cols <- c('HomeTeam', 'AwayTeam', 'B365H', 'B365D', 'B365A', 'FTR')
bets <- bets[, cols, with = F]

opta <- fread('Opta Predictions.csv')
dat <- merge(opta, bets, by = c("HomeTeam", "AwayTeam"))
dat <- dat[order(dat$Week), ]
dat$pSumOdds <- round(with(dat, 1/B365H + 1/B365D + 1/B365A), 4)
```

Observation: We can make a risk-free return on any odds distribution with implied probability < 1
```{r}
iprob <- 0.99

x <- runif(3)
p <- x/sum(x)*iprob
odds <- 1/p
weights <- round(p/sum(p)*100, 2)

round(odds*weights, 2) # risk-free return
```

Some odds conversions (American/European)
```{r}
o2p <- function(x) # American odds
{
  for (i in 1:length(x))
  {
    x[i] <- ifelse(x[i]<0, x[i]/(x[i]-100), 100/(x[i]+100))
  }
  return(c(x/sum(x), sum(x)))
}

am2eu <- function(x)
{
  for (i in 1:length(x))
  {
    x[i] <- ifelse(x[i]<0, (x[i]-100)/x[i], (x[i]+100)/100)
  }
  return(x)
}

eu2am <- function(x)
{
  for (i in 1:length(x))
  {
    x[i] <- ifelse(x[i]<2, -100/(x[i]-1), (x[i]-1)*100)
  }
  return(x)
}
```


Place risky-free bets - idea of arbitrage betting
```{r}
place_bets <- function(x, total_bet = 100, max_bet=0, max_bet_idx=0, eu = F)
{
  if (eu) {p <- o2p(eu2am(x))}
  else {p <- o2p(x)}
  p_sum <- tail(p,1)
  if(p_sum>=1) 
  {
    output <- paste0("vig: ", sprintf("%.2f", (p_sum-1)*100), "%\n",
                     "NOT risk-free!", "\n\n")
    return(cat(output))
  }
  p <- head(p, length(x))
  odds <- 1/p/p_sum
  bets <- p*total_bet
  if (max_bet!=0)
  {
    ratio <- max_bet/bets[max_bet_idx]
    bets <- bets*ratio
  }
  output <- paste0("bets: $", paste(sprintf("%.2f",bets), collapse = "-$"), "\n", 
                   "total bets: $", sprintf("%.2f",sum(bets)), "\n", 
                   "risk-free return: $", sprintf("%.2f",mean(odds*bets)), "\n",
                   "risk-free profit: ", ifelse((odds*bets)[1]-sum(bets)<0, 
                                                sprintf("-$%.2f", abs((odds*bets)[1]-sum(bets))),
                                                sprintf("$%.2f", (odds*bets)[1]-sum(bets))),
                   "\n",
                   "profit margin: ", sprintf("%.2f",((odds*bets)[1]-sum(bets))/sum(bets)*100),
                   "%\n\n")
  return(cat(output))
}
```


```{r}
fd_epl <- c(110,220,240)
dk_epl <- c(110,225,250)
dk_epl <- c(120,190,250)

b365_cn <- c(-118, 300, 260) #c(-110, 300, 260)
fd_cn <- c(-120, 320, 280) #c(-110, 300, 250)
dk_cn <- c(-120, 330, 285) #c(-105, 300, 255)

place_bets(b365_cn[1:2])
place_bets(fd_cn[1:2])
place_bets(dk_cn[1:2])

place_bets(c(1.33, 5.5), eu = T)
```


Solving LP
  - w1 + w2 + w3 = 1 (weights of bets)
  - p1, p2, p3 are true (constant) probabilities
  - x1, x2, x3 are bookmaker's (constant) odds
  - objective: max expected profit
  - potentially: solve LP for n possible outcomes
  
1) Without risk tolerance
```{r}
p <- c(0.65, 0.2, 0.15) # true probability
x_prob <- c(0.77, 0.18, 0.1)
x <- 1/x_prob # odds by bookmaker

obj <- p*x
mat <- matrix(c(rep(1,3)), nrow = 1, byrow = TRUE)
rhs <- 1
constr <- "=="
lp_result <- lp("max", obj, mat, constr, rhs)

print(lp_result$solution)
print(lp_result$objval)
```
Unsurprisingly, the solution to this LP will always tell us to bet on the outcome where the true probability exceeds the bookmaker's probability by the greatest margin.


2) With risk tolerance 
- defined using true probability p: p*w >= [tolerance] 
(perhaps there's a better way to incorporate this??)
```{r}
p <- c(0.65, 0.2, 0.15)
x_prob <- c(0.77, 0.18, 0.1)
x <- 1/x_prob
tol <- 0.3 # risk tolerance

obj <- p*x
mat <- matrix(c(rep(1,3),
                p), nrow = 2, byrow = TRUE)
rhs <- c(1, tol)
constr <- c("==", ">=")
lp_result <- lp("max", obj, mat, constr, rhs)

print(lp_result$solution)
print(lp_result$objval)
```

Dirichlet distribution: $f(\theta_1, \theta_2, ..., \theta_k; \alpha_1, \alpha_2, ..., \alpha_k) = \frac{1}{B(\alpha)} \prod_{i=1}^{k} \theta_i^{\alpha_i - 1}$
    - would the alphas be the bookmaker's probability?
```{r}

# bets[HomeTeam=="Man United" & AwayTeam=="Liverpool"]
odds <- c(4.75, 4.5, 1.6)
alpha <- 1/odds/sum(1/odds) # normalize or no?
# gtools::ddirichlet(x=c(0.3,0.3,0.4), alpha)


n <- 100
h_prob <- seq(0, 1, length.out = n+1)
d_prob <- seq(0, 1, length.out = n+1)
plane <- expand.grid(h_prob, d_prob)
colnames(plane) <- c('H', 'D')
plane <- plane[plane$H + plane$D <= 1, ]
plane$A <- 1-rowSums(plane)
# plane[duplicated(plane), ]

plane$ER <- apply(plane, 1, function(row) sum(row^2*odds))
# plane[sample(1:nrow(plane), 5),]

plane$density <- apply(plane[,1:3], 1, function(row) ddirichlet(row, alpha))
```


```{r}
density <- log(sqrt(plane$density+1))

plot_ly() %>%
  add_trace(
    type = "scatter3d",
    mode = "lines",
    x = c(0, 1, 0),
    y = c(0, 0, 1),
    z = c(1, 0, 0),
    line = list(color = "black"),
    showlegend = FALSE
  ) %>%
  add_trace(
    type = "scatter3d",
    mode = "markers",
    x = plane$H,
    y = plane$D,
    z = plane$A,
    marker = list(
      color = density,
      colorscale = "Viridis",
      cmin = min(density),
      cmax = max(density),
      colorbar = list(title = "Density"),
      size = 1.5
    )
  ) %>%
  layout(
    scene = list(
      xaxis = list(title = "Home Win"),
      yaxis = list(title = "Draw"),
      zaxis = list(title = "Away Win")
    )
  )
```


```{r}
options(digits = 4)

colMeans(rdirichlet(5e4, c(.7,.2,.1)))
colMeans(rdirichlet(5e4, c(.7,.2,.1)*10))

var(rdirichlet(5e4, c(.7,.2,.1)))
var(rdirichlet(5e4, c(.7,.2,.1)*10))

options(default)
```

1) For a single game, multiply Opta by a positive constant (multiplier)
    - generate true probabilities using rdirichlet where alpha = Opta*multiplier
    - compute weight distribution following a set strategy (ex: optimize LP)
    - for each simulated probabilities, compute expected profit using bookmaker's odds
        > sum(weight*odds*trueprob)
```{r}
multiplier_list <- c(10, 100, 1000, 10000)
opta_prob <- c(0.55, 0.3, 0.15)
bm_prob <- c(0.65, 0.25, 0.1)
# bm_prob <- c(0.56, 0.295, 0.145)
bm_odds <- 1/bm_prob

par(mfrow = c(2,2))
for (multiplier in multiplier_list)
{
  alpha <- multiplier*opta_prob
  nsim <- 5e4
  true_prob <- rdirichlet(nsim, alpha)
  
  # Optimize LP (i.e. pick the outcome with greatest difference in prob)
  diff <- sweep(true_prob, 2, bm_prob) 
  # diff <- t(t(true_prob) - bm_prob)
  weight_idx <- apply(diff, 1, function(row) which.max(row))
  # return <- mapply(function(idx, row) bm_odds[idx] * row, weight_idx, true_prob)
  return <- rep(NA, nsim)
  for (i in 1:nsim)
  {
    idx <- weight_idx[i]
    return[i] <- bm_odds[idx]*true_prob[i,idx]
  }
  hist(return, freq = T, main = '', xlab = "E[Return]",
       xlim = c(1,6))
  legend("topright", bty = "n",
         legend = c(paste0('mu =', round(mean(return),3)), 
                    paste0('sd =', round(var(return),3))), 
         title = paste0('Multiplier =', multiplier),
         pch = 20,
         inset = c(0.3, 0), cex = 0.8,
         col = "black")
}

```

We can lower our risk (variance) by placing arbitrage bets; lower expected return as tradeoff
    - probably more useful when true prob deviates from bookmaker more?
```{r}
multiplier <- 100
opta_prob <- c(0.55, 0.3, 0.15)
bm_prob <- c(0.65, 0.25, 0.1)
# bm_prob <- c(0.56, 0.295, 0.145)
bm_odds <- 1/bm_prob
alpha <- multiplier*opta_prob
nsim <- 5e4
true_prob <- rdirichlet(nsim, alpha)

# Optimize LP (i.e. pick the outcome with greatest difference in prob)
diff <- sweep(true_prob, 2, bm_prob) 
# diff <- t(t(true_prob) - bm_prob)
# weight_mat <- apply(diff, 1, function(row) place_bets(5,eu=F))

compute_weights <- function(p, total_bet = 1, eu = F)
{
  odds <- 1/p/sum(p)
  bets <- p*total_bet
  return(bets)
}

# weight should be a length 3 vector now!
weight_idx <- t(apply(diff, 1, function(row) {row[which.min(row)] <- 0
                                              row[row!=0] <- 1
                                              return (row)}))
weights <- weight_idx*true_prob
weight_mat <- sweep(weights, 1, rowSums(weights), "/")
# weights <- weights/replicate(3, rowSums(weights))

return <- rep(NA, nsim)
for (i in 1:nsim)
{
  return[i] <- sum(weight_mat[i,]*bm_odds*true_prob[i,])
}
hist(return, freq = T, main = '', xlab = "E[Return]")
legend("topright", bty = "n",
       legend = c(paste0('mu =', round(mean(return),3)), 
                  paste0('sd =', round(var(return),3))), 
       title = paste0('Multiplier =', multiplier),
       pch = 20,
       inset = c(0.3, 0), cex = 0.8,
       col = "black")
```

Simulate game outcomes using dirchlet centered around Opta*multiplier
    - sample(.., prob = c())
    - this is for a sequence of games
    - compute the return from each game (could be 0)
 > What does this really add to our simulation besides some uncertainty? 
 > In the long run it will be the same as simulating just using opta probs
```{r}
outcomes <- c("H", "D", "A")

simulate <- function(multiplier = 1, nsim = 100)
{
  # pred <- matrix(NA, nrow(dat), nsim)
  
  # for (i in 1:nsim)
  # {
  #   pred[,i] <- apply(dat[, c("pH", "pD", "pA")], 1,
  #                     function(row) sample(outcomes, 1, prob = rdirichlet(1, row*multiplier)))
  # }

  pred <- replicate(nsim, apply(dat[, c("pH", "pD", "pA")], 1,
                                function(row) sample(outcomes, 1, 
                                                     prob = rdirichlet(1, row*multiplier))))
  return(pred)
}

start <- proc.time()
pred_mat <- simulate(1, 2e3)
print(proc.time()-start)
```

```{r}
er_mat <- matrix(NA, nrow(pred_mat), ncol(pred_mat)) # expected return

for(i in 1:nrow(pred_mat))
{
  winning_odds <- with(dat, ifelse(FTR[i] == "H", B365H[i], 
                                   ifelse(FTR[i] == "A", B365A[i], B365D[i])))
  er_mat[i,] <- ifelse(pred_mat[i,]==dat$FTR[i], winning_odds, 0)
}
# View(er_mat)

return <- colMeans(er_mat)
# summary(return)
hist(return, freq = T, breaks = 40, 
     main = '', xlab = "return per game", ylab = '')
legend("topright", bty = "n",
       legend = c(paste0('mu =', round(mean(return),3)),
                  paste0('mdn =', round(median(return),3)),
                  paste0('sd =', round(sd(return),3)),
                  paste0('p(x>1) =', round(mean(return>1),3))),
       title = paste0('Multiplier =', 100),
       pch = 20,
       inset = c(0.1, 0), cex = 0.8,
       col = "black")
```

compare the distribution/frequency of H/D/A
```{r}
pred_dist <- matrix(NA,ncol(pred),3)

for(i in 1:ncol(pred))
{
  pred_dist[i,] <- prop.table(table(pred[,i])[outcomes])
}

colors <- c("lightgreen", "skyblue", "salmon")
barplot(t(pred_dist), col = colors, 
        main = "", xlab = "", ylab = "P(X=x_i)", 
        border = NA, space = 0)
legend("top", legend = outcomes, fill = colors, 
       horiz = TRUE, xpd = TRUE, bty = "n")

round(prop.table(table(dat$FTR))[outcomes],4) # actual
cat((round(colMeans(pred_dist),4))) # simulation
```
Opta is overestimating draws - although it's true that in this season draws are less frequent than usual
